---
title: "Classifications and hierarchy trees"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Classifications
There are variables whose values can change over time. An example is the name
of a municipality. For these cases, there is the concept of classifications,
which allows access to the values of a specific classification. We can request
the available classifications with the `get_metadata_classifications()` function.

```{r classifications1}
library(ineapir)

# Get classifications
classifications <- get_metadata_classifications(lang = "EN")
head(classifications)
```

In the case of municipalities, there is a classification for each year, so we
can retrieve the municipalities for a particular year using the
`get_metadata_values()` function and the `classification` argument.

```{r classifications2}
# Select the classifications with name 'Geographical yyyy'
classifications <- get_metadata_classifications(lang = "EN")
head(classifications[grepl("geographical", classifications$Nombre,
                           ignore.case = TRUE),])

# Municipalities: id=19
# To retrieve the municipalities of 2007 we use the classificacion with id=29
municipalities <- get_metadata_values(variable = 19, classification = 29)
head(municipalities)

```

## Hierarchy trees
There are certain values that belong to a hierarchical structure and can
have parents and children. To obtain the children of a specific value, we 
can use the `get_metadata_values()` function with the `variable` and `value`
arguments. For example, if we want to find the provinces of Galicia:
```{r tree1}
# Variable: Autonomous communities (id=70)
# Value: Galicia (id=9008)
# Get the children of id=9008 (provinces of Galicia)
provinces <- get_metadata_values(variable = 70, value = 9008)
provinces
```

If we want to go deeper into the hierarchical structure we can use the
`hierarchy` argument, which represents the depth.

```{r tree2}
# Variable: Autonomous communities (id=70)
# Value: Galicia (id=9008)
# Get the children of each province (municipalities of Galicia)
municipalities <- get_metadata_values(variable = 70, value = 9008, hierarchy = 1)
head(municipalities)
```


If we want the root of the tree to be the variable, we do not specify any value.
```{r tree3}
# Variable: Autonomous communities (id=70)
# Get the children of each Autonomous communities (provinces)
provinces <- get_metadata_values(variable = 70, hierarchy = 1)
head(provinces)
```


Additionally, we can filter out the variables and values that only interest
us with the `filter` argument.


1. Example 1.
```{r tree4}
# We define the filter as a list of variables and values
filter <- list("70" = 9008 # variable id = 70, value id = 9008 (Galicia)
               )
# Get the children of id=9008 (provinces of Galicia)
provinces <- get_metadata_values(variable = 70, filter = filter, hierarchy = 1,
                                 validate = FALSE)
head(provinces)
```


2. Example 2.
```{r tree5}
# We define the filter as a list of variables and values
filter <- list("115" =  "" # variable id = 115, all values
               )
# Get the children of id=70 (provinces of Spain)
provinces <- get_metadata_values(variable = 70, filter = filter, hierarchy = 1,
                                 validate = FALSE)
head(provinces)
```


3. Example 3.
```{r tree6}
# We define the filter as a list of variables and values
filter <- list("70" = 9008, # variable id = 70, value id = 9008 (Galicia) 
               "115" = 28 # variable id = 115, value id = 28 (Lugo)
               )
# Get the children of id=28 (municipalities of Lugo province)
municipalities <- get_metadata_values(variable = 70, filter = filter, 
                                      hierarchy = 2, validate = FALSE)
head(municipalities)
```

4. Example 4.
```{r tree7}
# We define the filter as a list of variables and values
filter <- list("70" = 9008, # variable id = 70, value id = 9008 (Galicia)  
               "115" =  28 , # variable id = 115, value id = 28 (Lugo)
               "19" = 570 # variable id = 19, value id = 570 (Lugo)
               )
# Get the children of id=570 (census sections of Lugo municipality)
sections <- get_metadata_values(variable = 70, filter = filter, hierarchy = 4,
                                validate = FALSE)
head(sections)
```

