---
title: "How to filter data using shortcut names"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Instead of using numeric codes, it is possible to use certain shortcut names
when the `filter` argument is passed. However, for a better performance is
recommended to use numeric codes. The available shortcuts for variables are
shown below.

|Shortcut | Variable ID | Comment|
|---------|----|--------|
|nac             | 349 | National|
|ccaa            | 70  | Autonomous Communities |
|prov            | 115 | Provinces |
|mun             | 19  | Municipalities |
|island          | 20  | Islands |
|isla            | 20  | Islands |
|group           | 762 | CPI groups |
|grupo           | 762 | CPI groups |
|subgroup        | 763 | CPI subgroups |
|subgrupo        | 763 | CPI subgroups |
|class           | 764 | CPI class |
|clase           | 764 | CPI class |
|subclass        | 765 | CPI subclass |
|subclase        | 765 | CPI subclass |
|heading         | 270 | CPI headings |
|rubrica         | 270 | CPI headings |
|specialgroup    | 269 | CPI special groups |
|grupoespecial   | 269 | CPI special groups |
|datatype        | 3   | Type of data |
|tipodato        | 3   | Type of data |
|sex             | 18  | Sex |
|sexo            | 18  | Sex |
|age1            | 355 | Simple age |
|edad1           | 355 | Simple age |
|aget            | 356 | Age totals |
|edadt           | 356 | Age totals |
|ageg            | 360 | Age groups |
|edadg           | 360 | Age groups |
|ages            | 357 | Age semi-intervals |
|edads           | 357 | Age semi-intervals |
|age             |355, 356, 357, 360 | Age wrapper |
|edad            |355, 356, 357, 360 | Age wrapper |
|generation      | 612 | Generation/ages |
|generacion      | 612 | Generation/ages |
|nationality     | 141 | Nationality |
|nacionalidad    | 141 | Nationality |
|birthcountry    | 431, 432 | Country of birth |
|paisnacimiento  | 431, 432 | Country of birth |
|birthplace      | 97  | Place of birth |
|lugarnacimiento | 97  | Place of birth |
|effectscorr     | 544 | Correction of effects |
|efectoscorr     | 544 | Correction of effects |
|values          | -    | Values wrapper |

## Filtering data from tables

To use a filter with shortcuts it is necessary to pass the argument
`filter` using a shortcut as variable and a name as value.

### Example one
We are going to filter the table [Índices nacionales: general y de grupos ECOICOP](https://www.ine.es/jaxiT3/Tabla.htm?t=50902) (Id 50902):

```{r tabla_filter_1}
library(ineapir)

# Filter using shortcuts (CPI)
filter <- list(tipodato = "variación anual" , # variable id = 3
               grupo = "índice general"       # variable id = 762
               )

# Request data using the filter
ipc <- get_data_table(idTable = 50902, filter = filter, unnest = TRUE,
                      tip = "A", nlast = 5, validate = FALSE)
ipc[,c("Nombre", "T3_Periodo", "Anyo", "Valor")]
```

Besides, we can also use the *values* wrapper.

```{r tabla_filter_2}
# Filter with the values wrapper
filter <- list(values = c("variación anual" , "índice general"))

# Request data using the filter
ipc <- get_data_table(idTable = 50902, filter = filter, unnest = TRUE,
                      tip = "A", nlast = 5, validate = FALSE)
ipc[,c("Nombre", "T3_Periodo", "Anyo", "Valor")]
```
If we add a minus sign to a value, the value will be excluded from the filter.

```{r tabla_filter_2b}
# Filter with the values wrapper
filter <- list(values = c("variación anual" , "-índice general"))

# Request data using the filter
ipc <- get_data_table(idTable = 50902, filter = filter, unnest = TRUE,
                      tip = "A", nlast = 2, validate = FALSE)
ipc[,c("Nombre", "T3_Periodo", "Anyo", "Valor")]
```
### Example two
We are going to filter the table [PIB pm Oferta (Indices de volumen encadenado)](https://www.ine.es/jaxiT3/Tabla.htm?t=30679) (Id 30679):

```{r tabla_filter_3}
# Filter with the values wrapper (GDP)
filter <- list(values = c("producto interior", "datos ajustados", "variación anual"))

# Request data using the filter and shortcut = TRUE
pib <- get_data_table(idTable = 30679, filter = filter, unnest = TRUE,
                      tip = "A", nlast = 5, validate = FALSE)
head(pib[,c("Nombre", "T3_Periodo", "Anyo", "Valor")])
```

### Example three
We are going to filter the table [Población residente por fecha, sexo y edad (desde 1971)](https://www.ine.es/jaxiT3/Tabla.htm?t=56934) (Id 56934)

```{r tabla_filter_4}
# Filter with the values wrapper (population)
filter <- list(values = c("todas las edades", "total"))

# Request data using the filter
pob <- get_data_table(idTable = 56934, filter = filter, unnest = TRUE, 
                      tip = "A", nlast = 5, validate = FALSE)
pob[,c("Nombre", "T3_Periodo", "Anyo", "Valor")]
```

## Filtering data from series
It is necessary to pass the argument `filter` of the 
`get_data_series_filter()` function. We use a shortcut as variable 
and a name as value.

### Example one

```{r series_filter_1}
# Filter with the values wrapper (CPI)
filter <- list(values = c("variación anual" , "índice general", "total nacional"))

# Request data using the filter
ipc <- get_data_series_filter(operation = "IPC", filter = filter,
                              periodicity = 1, unnest = TRUE, tip = "A",
                              validate = FALSE)
ipc[,c("Nombre", "T3_Periodo", "Anyo", "Valor")]
```

### Example two

```{r series_filter_2}
# Filter with the values wrapper (GDP)
filter <- list(values = c("producto interior", "datos ajustados",
                       "volumen encadenado", "variación anual"))

# Request data using the filter
pib <- get_data_series_filter(operation = "CNTR2010", filter = filter,
                              periodicity = 3, unnest = TRUE, tip = "A",
                              validate = FALSE)
pib[,c("Nombre", "T3_Periodo", "Anyo", "Valor")]
```


### Example three

```{r series_filter_3}
# Filter with the values wrapper combine with individual shortcuts (population)
filter <- list(values = c("todas las edades", "total nacional"), nacionalidad = "total",
               sexo = "total", paisnacimiento = "total" )

# Request data using the filter
pob <- get_data_series_filter(operation = "ECP", filter = filter,
                              periodicity = 3, unnest = TRUE, tip = "A",
                              validate = FALSE)
pob[,c("Nombre", "T3_Periodo", "Anyo", "Valor")]
```
