---
title: "Tips"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Use of `verbose`

When the argument `verbose = TRUE` is passed to a function, a number of 
information is displayed in the console, including the URL used to call the API.

```{r verbose}
library(ineapir)

# Variables used in the operation IPC and verbose = TRUE
variables <- get_metadata_variables(operation = "IPC", verbose = TRUE)
```

## Use of `validate`
If it is not necessary to validate the values of the arguments passed to a
function, this validation can be turned off with `validate = FALSE`.

```{r validate_1}
# Variables used in the operation IPC with verbose = TRUE and validate = FALSE
variables <- get_metadata_variables(operation = "IPC", verbose = TRUE,
                                    validate = FALSE)
```

This is useful when we are filtering data and the filter we are using is
the one we want. Turning off validation means making fewer API calls.
For better performance, it is recommended to use codes in the filter instead
of shortcuts and disable validation if we are sure that the filter is correct.

```{r validate_2}
# The filter 
filter <- list("115" = "2",     # variable id = 115, value id = 2
               "3" = "74" ,     # variable id = 3, value id = 74
               "762" = "304092" # variable id = 762, value id = 304092
               )

# Request data using the filter with verbose = TRUE and validate = TRUE
ipc <- get_data_series_filter(operation = "IPC", filter = filter, periodicity = 1,
                       verbose = TRUE)

# Request data using the filter with verbose = TRUE and validate = FALSE
ipc <- get_data_series_filter(operation = "IPC", filter = filter, periodicity = 1,
                       verbose = TRUE, validate = FALSE)
```

## Use of `lang`
It is possible to retrieve information in English passing the argument
`lang = "EN"` to certain functions.

```{r lang_1}
# Variables used in the operation IPC and lang = "EN"
variables <- get_metadata_variables(operation = "IPC", lang = "EN")
variables
```


We can use the English language as well filtering data using shortcuts.

```{r lang_2}
# Filter with the values wrapper (CPI)
filter <- list(values = c("annual change" , "overall index"))

# Request data using the filter with shortcut = TRUE and lang = "EN"
cpi <- get_data_table(idTable = 50902, filter = filter, unnest = TRUE,
                      tip = "A", nlast = 5, lang = "EN")

cpi[,c("Nombre", "T3_Periodo", "Anyo", "Valor")]
```

## Use of `unnest`
When we request data from tables or series using the functions
`get_data_table()`, `get_data_series()` and `get_data_series_filter`, the
`$Data` column containing the data values is a list of data frames, one for each
 entry.

```{r unnest_1}
# Filter with the values wrapper (CPI)
filter <- list(values = c("variación anual" , "variación mensual", "índice general"))

# Request data using the filter and unnest = FALSE
ipc <- get_data_table(idTable = 50902, filter = filter, tip = "A", nlast = 3)

# We have two entries
 ipc[,c("COD","Nombre")]
 
# Data column is a list of two data frames. One for each entry. 
 ipc$Data
```

If we want to get only one data frame for all data values, we can pass
the argument `unnest = TRUE`.

```{r unnest_2}
# Filter with the values wrapper (CPI)
filter <- list(values = c("variación anual" , "variación mensual", "índice general"))

# Request data using the filter and unnest = TRUE
ipc <- get_data_table(idTable = 50902, filter = filter, tip = "A",
                      nlast = 3, unnest = TRUE)

# Unique data frame with 6 rows
ipc[,c("Nombre", "T3_Periodo", "Anyo", "Valor")]
```

## Use of `metanames` and `metacodes`
When we pass the argument `metanames = TRUE` we get additional columns that 
contain the names of the values extracted from the metadata information.
To get additional columns containing the codes and ids of the values 
extracted from the metadata information we pass the argument `metacodes = TRUE`.

```{r extractmetadata}
# Filter for total population of men and women 
filter <- list(nacionalidad = "total", sexo = c("hombres", "mujeres"))

# Table: Resident population by date, sex, age group and nationality
poblacion <- get_data_table(idTable = 9674, filter = filter, validate = FALSE, 
                            tip = "AM", unnest = TRUE, nlast = 1)
# Names of columns
names(poblacion)

# Request data with metanames = TRUE
poblacion <- get_data_table(idTable = 9674, filter = filter, validate = FALSE, 
                            tip = "AM", unnest = TRUE, nlast = 1,
                            metanames =  TRUE, metacodes = TRUE)

# Additional columns of Nacionalidad, Edad and Sexo
names(poblacion)
```

## Filtering with regular expresions
When we use shortcuts names in the filter, we can also include as values
regular expressions if this helps us filter the data better.

```{r regexp_1}
# Filter with a regular expression for ages between 20 and 25 years old
filter <- list(sexo = "total", edad = "2[0-5]+")

# Request data using the filter with a regular expression
pob <- get_data_table(idTable = 56934, filter = filter, tip = "A", nlast = 1,
                      unnest = TRUE)

# Only ages between 20 and 25 years old
pob[,c("Nombre", "Anyo", "Valor")]
```

